From Reversible Programs to \\ Univalent Universes and Back

### 2017-06-30 upi.tex
move the wavefront in section 3

@@

Fig.~\ref{fig:fib} depicts a type family $P : A \rightarrow \mathcal{U}$
mapping each member $x : A$ to a type $P(x)$. As everything is continuous,
we expect this type family to
respect identities. In other words if $x:A$ and $y:A$ are identified using a
path $x \equiv y$, it must
be the case that the spaces $P(x)$ and $P(y)$ are equivalent
$P(x) \simeq P(y)$. Generally speaking, there is no reason to
expect that the spaces $x \equiv y$ and $P(x) \simeq P(y)$ to be themselves
equivalent for all $x$ and $y$. And indeed they are most frequently not
equivalent. In the rare cases in which they are equivalent, the fibration is a
\emph{univalent fibration}.

\begin{figure}
\begin{tabular}{c@{\qquad\qquad}c}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $A$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$x$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$y$};
  \draw (-1,-2) ellipse (0.5cm and 2cm);
  \node[left] at (-1.5,-2) {Fiber $P(x)$};
  \draw (1,-2) ellipse (0.5cm and 2cm);
  \node[right] at (1.5,-2) {Fiber $P(y)$};
\end{tikzpicture}
&
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $A$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$x$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$y$};
  \draw (-1.3,-2) ellipse (0.5cm and 2cm);
  \node[left] at (-1.8,-2) {Fiber $P(x)$};
  \draw (1.3,-2) ellipse (0.5cm and 2cm);
  \node[right] at (1.8,-2) {Fiber $P(y)$};
  \draw[below,cyan,thick] (-1,-5) -- (1,-5);
  \node[below,cyan,thick] at (0,-5) {$\equiv$};
  \draw[->,red,ultra thick] (-0.8,-1.7) to [out=45, in=135] (0.8,-1.7);
  \draw[->,red,ultra thick] (0.8,-2.3) to [out=-135, in=-45] (-0.8,-2.3);
  \node[red,ultra thick] at (0,-2) {$\simeq$};
\end{tikzpicture}
\end{tabular}
\caption{\label{fig:fib}(left) Type family $P : A \rightarrow \mathcal{U}$ as a
  fibration with total space $\Sigma_{(x:A)} P(x)$; (right) a path $x \equiv y$
  in the base space induces an equivalence between the spaces $P(x)$ and $P(y)$}
\end{figure}

>>>>

Fig.~\ref{fig:fib} on the left depicts a type family $P : A \rightarrow \mathcal{U}$
mapping each member $x : A$ to a type $P(x)$. As everything is continuous,
we expect this type family to respect identities. In other words if $x:A$ and
$y:A$ are identified using a path $x \equiv y$ (depicted on the right), it must
be the case that the spaces $P(x)$ and $P(y)$ are equivalent
$P(x) \simeq P(y)$. Although we depict the path ``in'' the Base Space $A$
(and that is indeed the case in classical homotopy theory), it is best to
think of the evidence given by $\equiv$ as being outside $A$ (but in $\mathcal{U}$).

Generally speaking, there is no reason to
expect that the spaces $x \equiv y$ and $P(x) \simeq P(y)$ of
``evidence'' that identify $x$ and $y$, and $P(x)$ and $P(y)$ respectively,
are themselves related, never mind equivalent, for all or even any $x$ and $y$.
In the rare cases in which these ``spaces of evidence'' are equivalent, such a
fibration is a \emph{univalent fibration}.

\begin{figure}
\begin{tabular}{c@{\qquad\qquad}c}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $A$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$x$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$y$};
  \draw (-1,-2) ellipse (0.5cm and 2cm);
  \node[left] at (-1.5,-2) {Fiber $P(x)$};
  \draw (1,-2) ellipse (0.5cm and 2cm);
  \node[right] at (1.5,-2) {Fiber $P(y)$};
\end{tikzpicture}
&
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $A$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$x$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$y$};
  \draw (-1.3,-2) ellipse (0.5cm and 2cm);
  \node[left] at (-1.8,-2) {Fiber $P(x)$};
  \draw (1.3,-2) ellipse (0.5cm and 2cm);
  \node[right] at (1.8,-2) {Fiber $P(y)$};
  \draw[below,cyan,thick] (-1,-5) -- (1,-5);
  \node[below,cyan,thick] at (0,-5) {$\equiv$};
  \draw[->,red,ultra thick] (-0.8,-1.7) to [out=45, in=135] (0.8,-1.7);
  \draw[->,red,ultra thick] (0.8,-2.3) to [out=-135, in=-45] (-0.8,-2.3);
  \node[red,ultra thick] at (0,-2) {$\simeq$};
\end{tikzpicture}
\end{tabular}
\caption{\label{fig:fib}(left) Type family $P : A \rightarrow \mathcal{U}$ as a
  fibration with total space $\Sigma_{(x:A)} P(x)$;\newline
 (right) a path $x \equiv y$
  in the base space induces an equivalence between the spaces $P(x)$ and $P(y)$}
\end{figure}

@@
